Viral Sequence Sorting Tools - Rules on Fragments

### 2023-05-26 TestingPipeline/Scripts/paper_figure_generation_balanced_testing.Rmd
Comparing 50:50 vs 90:10 not viral:viral splits.

@@

## Calculate the performance of each pipeline
quantifying variability between rule sets
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*8),
                      precision=rep(0, nrow(combos_list)*8),
                       recall=rep(0, nrow(combos_list)*8),
                       F1=rep(0, nrow(combos_list)*8),
                       MCC=rep(0, nrow(combos_list)*8), 
                      AUC=rep(0, nrow(combos_list)*8))

accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==1,], combos_list))
for (i in 2:8) {
  accuracy_scores <- rbind(accuracy_scores,
                           cbind(testing_set_index=rep(i, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==i,], combos_list)))
}
```

@@

@@

@@

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(viridis(6)), 1))
```

@@

@@

@@

@@

```{r}
pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
pal_orange <- RColorBrewer::brewer.pal(1, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange[1]), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free")
```

@@

viruses$keep_score_all_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

viruses$keep_score_all_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

viruses$keep_score_all_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

@@

@@

@@

pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")

@@

ann_colors = list(
  tnv = c("black", pal(6)[1]),
  tv = c("black", pal(6)[3]),
  dvf = c("black", pal(6)[2]),
  vb = c("black", pal(6)[4]),
  vs = c("black", pal(6)[5]),
  vs2 = c("black", pal(6)[6]),
  score = c("archaea"=pal(6)[1], "bacteria"=pal(6)[2], "fungi"=pal(6)[3],
              "plasmid"=pal(6)[4], "protist"=pal(6)[5], "virus"=pal(6)[6])
)

@@

pheatmap::pheatmap(t(viral_scores_1),
                  annotation_col = ann_row,
                  annotation_row = ann_col,
                  annotation_colors = ann_colors,
                  #border_color = "black",
                  color = c(rev(RColorBrewer::brewer.pal(3,"Oranges")),
                            RColorBrewer::brewer.pal(5,"Purples") 
                            ),
                  show_rownames = F,
                  show_colnames = F,
                  annotation_legend = F,
                  cluster_rows = F,
                  cluster_cols = F,
                  cellheight=3,
                  filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores.png"
                  )
```

>>>>

only 5 testing sets   
```{r}
viruses <- viruses %>% filter(Index<=5)
```

balancing testing set
1000 viruses per testing set
600 bacteria
100 arachaea
100 fungi
100 plasmid
100 protist
```{r}
viruses <- rbind(viruses %>% filter(seqtype=="archaea") %>%
  group_by(Index) %>%
  slice_sample(n=100),
  viruses %>% filter(seqtype=="bacteria") %>%
  group_by(Index) %>%
  slice_sample(n=600),
  viruses %>% filter(seqtype=="fungi") %>%
  group_by(Index) %>%
  slice_sample(n=100),
  viruses %>% filter(seqtype=="plasmid") %>%
  group_by(Index) %>%
  slice_sample(n=100),
  viruses %>% filter(seqtype=="protist") %>%
  group_by(Index) %>%
  slice_sample(n=100),
  viruses %>% filter(seqtype=="virus") %>%
  group_by(Index) %>%
  slice_sample(n=1000)
)

@@

## Calculate the performance of each pipeline
quantifying variability between rule sets
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*5),
                      precision=rep(0, nrow(combos_list)*5),
                       recall=rep(0, nrow(combos_list)*5),
                       F1=rep(0, nrow(combos_list)*5),
                       MCC=rep(0, nrow(combos_list)*5), 
                      AUC=rep(0, nrow(combos_list)*5))

accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==1,], combos_list))
for (i in 2:5) {
  accuracy_scores <- rbind(accuracy_scores,
                           cbind(testing_set_index=rep(i, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==i,], combos_list)))
}
```

@@

getting set of tools with roughly equivalent precision
```{r}
t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 1 0 0"])

t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 0 1 0 0"])

t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 0 0 0 1"])

t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 1 0 1"])
```
accounting for multiple t-tests with Bonferroni, then the following have the same precision:
- vs2+tnv, vb+tnv, vb, and vs2

@@

getting set of tools with roughly equivalent recall
```{r}
t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
       accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 1 1 1"])

t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
       accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 0 1"])

t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
       accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 1 0 1"])

t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
       accuracy_scores$recall[accuracy_scores$toolcombo=="1 0 0 0 1 1"])

t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
       accuracy_scores$recall[accuracy_scores$toolcombo=="1 0 0 1 1 1"])

t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
       accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 1 0 1 1"])
```
accounting for multiple t-tests with Bonferroni, then the following have the same recall:
- vs2+vs+dvf+tv, vs2+vs+dvf+tv+vb, vs2+dvf+tv, vs2+vs+dvf+tv, vs2+dvf+tv+vb,

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 1)) + 
  ylim(c(0,1))
```

```{r}

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(viridis(6)), 1))
```
```{r}
max(accuracy_scores$MCC)
```
=78

getting set of tools with roughly equivalent MCC
```{r}
t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$MCC[accuracy_scores$toolcombo=="0 1 1 0 0 1"])

t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"])

t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 0 0 0 1"])

t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 1 0 1"])

t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 1 1"])

t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 1 1 0 0 1"])
```
accounting for multiple t-tests with Bonferroni, then the following have the same MCC:
- vs2+tnv, vs2+tnv+dvf, vs2, vs2+tnv+tv+vb, vs2+tnv+tv+vs

@@

ggplot(accuracy_scores_melt[,], aes(x=numrules, y=performance_metric_score, 
                                  color=numrules, fill=numrules)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Score") +
  xlab("Number of Rule Sets") +
  scale_fill_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 1)) +
  facet_wrap(~performance_metric)

@@

```{r}
viruses$keep_score_all_tv_only_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = T,
                                              tv_1=F, tv_2=F, tv_3=F, tv_4=T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_all_tv_only_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = T,
                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_all_tv_only_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = T,
                                              tv_1=F, tv_2=T, tv_3=F, tv_4=F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_all_tv_only_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = T,
                                              tv_1=F, tv_2=F, tv_3=T, tv_4=F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_all_tv_only"))

viruses_long_scores <- apply(viruses_long_scores, c(1,2), function(x) {if (x>0){x<-1} else{x<-0}})

viruses_long_scores <- as.data.frame(cbind(viruses$true_virus, rowSums(viruses_long_scores)))

table(viruses_long_scores$V1, viruses_long_scores$V2>1)
```

@@

 viruses$keep_score_vs2_tv_ntv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

@@

```{r}
pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free")
```

@@

viruses$keep_score_all_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              ntv_1=F, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

viruses$keep_score_all_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              ntv_1=F, ntv_2=F, ntv_3=T, ntv_4=F, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

viruses$keep_score_all_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=T, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

viruses$keep_score_all_ntv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

@@

```{r}
table(viruses$keep_score_all_ntv_0, viruses$keep_score_all_ntv_1)
```

```{r}
table(viruses$keep_score_all_ntv_4, viruses$keep_score_all_ntv_5)
```

@@

```{r}
table(viruses$true_virus, viruses$keep_score_ind_dvf)
table(viruses$true_virus, viruses$keep_score_ind_vs2)
table(viruses$true_virus, viruses$keep_score_ind_vs)
table(viruses$true_virus, viruses$keep_score_ind_vb)
```
seems like false positive rate is way high for 0.5 across all

```{r}
table(viruses$true_virus, viruses$keep_score_ind_dvf+viruses$keep_score_ind_vs2)
```

```{r}
table(viruses$true_virus, viruses$keep_score_ind_dvf+viruses$keep_score_ind_vs2+viruses$keep_score_ind_ntv)
```

```{r}
table(viruses$keep_score_ind_vs2, viruses$keep_score_ind_dvf+viruses$keep_score_ind_vs2+viruses$keep_score_ind_ntv)
```
with 0.5 rules: 1464 sequences predicted viral that weren't viral by vs2 on its own
without them: 1234 sequences added

```{r}
table(viruses$keep_score_ind_vs2, viruses$keep_score_ind_dvf+viruses$keep_score_ind_vs2+viruses$keep_score_ind_ntv,
      viruses$true_virus)
```
with 0.5 rules: however, of these, only 432 (30%) are truly viruses
without 0.5 rules: only 375 (30%) are truly viruses

```{r}
table(viruses$keep_score_ind_vs2, viruses$keep_score_ind_vb+viruses$keep_score_ind_vs2+viruses$keep_score_ind_ntv,
      viruses$true_virus)
```
with 0.5 rules: however, of these, only 74 (7%) are truly viruses
without 0.5 rules: only 87 (4%) are truly viruses

@@

pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
pal2 <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")

@@

ann_colors = list(
  tnv = c("black", pal(6)[1]),
  tv = c("black", pal(6)[3]),
  dvf = c("black", pal(6)[2]),
  vb = c("black", pal(6)[4]),
  vs = c("black", pal(6)[5]),
  vs2 = c("black", pal(6)[6]),
  score = c("archaea"=pal2(20)[13], "bacteria"=pal2(20)[9], "fungi"=pal2(20)[4],
              "plasmid"=pal2(20)[19], "protist"=pal2(20)[5], "virus"=pal2(20)[17])
)

@@

pheatmap::pheatmap(t(viral_scores_1),
                  annotation_col = ann_row,
                  annotation_row = ann_col,
                  annotation_colors = ann_colors,
                  #border_color = "black",
                  color = c(rev(RColorBrewer::brewer.pal(3,"Oranges")),
                            RColorBrewer::brewer.pal(5,"Purples") 
                            ),
                  show_rownames = F,
                  show_colnames = F,
                  annotation_legend = F,
                  cluster_rows = F,
                  cluster_cols = F,
                  cellheight=3,
                  filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores_balanced.png"
                  )
```

### 2023-05-26 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
Comparing 50:50 vs 90:10 not viral:viral splits.

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
                                include_tuning_not_viral=FALSE,
                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
                                include_virsorter=FALSE) {

>>>>

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
                                include_tuning_not_viral=FALSE,
                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=T,
                                include_virsorter=FALSE) {
